Model Estimation

### 2020-05-11 05_estimation.Rmd
Clean up file names and headers
Re-estimate models and update discussion
Includes analysis of alternative distance calculations, addressing #5.

---
title: "Model "
author: "Greg Macfarlane"
date: "6/17/2019"
output: html_document
---

@@

```{r bg_access}
bg_access <- read_rds("data/bg_access.rds")
```

The accessibility is calculated at the block group level, but the other information
(health, demographics) is available only at the tract level.

```{r tracts}
mydata <- bg_access %>% 
  dplyr::mutate(tract = substr(geoid, 0, 11)) %>%
  left_join(
    read_rds("data/tracts.rds") %>%
      st_set_geometry(NULL),
    by = c("tract" = "geoid")
  ) %>%
  filter(!is.na(households))
```

### Contiguity weights

@@

```{r Wcontig}
areas_adj <- poly2nb(mydata)
Wcontig <- nb2listw(areas_adj, style = "W", zero.policy = TRUE)
print(Wcontig, zero.policy = TRUE)
```

@@

$$PhysAct = X\beta + \beta_{access}*access + u; u = \lambda W u + \epsilon$$
We will consider both spatial weights matrices.

```{r ols_vs_sem}
access_ols <- lm(update(base_formula, physact ~ . + access_ls), data = mydata)
access_sem_wc <- errorsarlm(update(base_formula, physact ~ . + access_ls),
                            data = mydata, listw = Wcontig, zero.policy = TRUE)

ols_vs_sem <- list("OLS" = access_ols, "SEM" = access_sem_wc)
```

@@

sp_hausman(access_ols, access_sem_wc)
```

@@

I'm curious what happens when we remove density from the base model.

```{r nodensity}
# take log(density) out of the bare access model
nodensity <- lapply(ols_vs_sem, function(model) {
  update(model, . ~ . - log(density))
})
lapply(nodensity, tidy) %>%
  bind_rows(.id = "model") %>%
  dwplot() + scale_x_continuous(limits = c(-0.3, 1))
```

Basically the same thing. It's worth looking at whether the same thing is happening
in the obesity model.

```{r obesity_sem}
obesity_sem <- lapply(ols_vs_sem, function(model) {
  update(model, obesity ~ . + physact)
})
lapply(obesity_sem, tidy) %>%
  bind_rows(.id = "model") %>%
  dwplot() 
```

Yup. Okay, given that, the question is whether adding lagged
variables $\gamma W X$ to the model will resolve the discrepancy.

@@

The same basic logic applies to the SLX and SDEM models; if there is no dependence
relationship among the $y$ variables, then the SLX and SDEM will have unbiased
estimates of $\beta$ and $\gamma$.

```{r access_slx}
access_slx_wc <- lmSLX(update(base_formula, physact ~ . + access_ls), 
                       data = mydata, listw = Wcontig, zero.policy = TRUE)
access_sdem_wc <- errorsarlm(update(base_formula, physact ~ . + access_ls),
                            data = mydata, listw = Wcontig, zero.policy = TRUE, 
                            Durbin = TRUE)
slx_vs_sdem <- list(
  "SDEM" = access_sdem_wc,
  "SLX" = access_slx_wc
)
```

As before, we can try a Hausman-style test.
```{r slx_hausman}
slx_hausman <- function(slx, sdem) {
  b1 <- slx$coefficients
  b2 <- sdem$coefficients
  db <- b1 - b2

@@

  tibble(
    Hstat = t(db) %*% MASS::ginv(dV) %*% db, 
    "p-value" = pchisq(Hstat, df = Matrix::rankMatrix(dV), lower.tail = FALSE)
  )
}
slx_hausman(access_slx_wc, access_sdem_wc)
```

@@

```{r lagged_vars}
contiglagged_vars <- mydata %>%
  as_tibble() %>%
  transmute_at(
    vars(density, income, fulltime, college, single, youth, young_adults, seniors, black, asian, hispanic, other, physact),
    ~lag.listw(Wcontig, .)
  ) %>%
  mutate(geoid = mydata$geoid) %>%
  dplyr::select(geoid, everything())

names(contiglagged_vars)[-1] <- str_c("lag.", names(contiglagged_vars)[-1])
contiglagged_data <- mydata %>% left_join(contiglagged_vars, by = "geoid")
```

@@

```{r nolag_access}
access_slx_wc <- lm(update(
  base_formula, physact ~ . +  log(lag.density) + log(lag.income) + 
    lag.fulltime + lag.college + lag.single + lag.youth +  lag.young_adults + 
    lag.seniors + lag.black + lag.asian + lag.hispanic + lag.other + access_ls),
  data = contiglagged_data)

access_sdem_wc <- errorsarlm(update(
  base_formula, physact ~ . +  log(lag.density) + log(lag.income) + 
    lag.fulltime + lag.college + lag.single + lag.youth +  lag.young_adults + 
    lag.seniors + lag.black + lag.asian + lag.hispanic + lag.other + access_ls), 
  data = contiglagged_data, listw = Wcontig, zero.policy = TRUE)
```

```{r nolag_access_tests}
slx_hausman(access_slx_wc, access_sdem_wc)
slx1_vs_sdem1 <- list("SLX" = access_slx_wc, "SDEM" = access_sdem_wc)

screenreg(slx1_vs_sdem1, ci.force = TRUE)
```

@@

```{r estimate_sdems}
base_sdem <- update(access_sdem_wc, formula = . ~ . - access_ls,
                    data = contiglagged_data, listw = Wcontig, zero.policy = TRUE)

# estimate models
pa_access_sdm <- access_sdem_wc
pa_multi_sdm   <- update(base_sdem, .~ . + multi_ls)
```

@@

```{r access_sdm}
pa_models <- list("Base"        = base_sdem,
                  "Logsum"      = pa_access_sdm,
                  "Attributes"  = pa_multi_sdm)
```

>>>>

---
title: "Model Estimation"
---

@@

```{r tracts}
est_data <- read_rds("data/tracts.rds") %>%
  left_join(read_rds("data/tract_logsums.rds"), by = "geoid") %>%
  left_join(read_rds("data/walk10.rds"), by = "geoid")
```

@@

```{r W}
Wtr <- nb2listw(poly2nb(est_data), style = "W", zero.policy = TRUE)
print(Wtr, zero.policy = TRUE)
```

@@

$$PhysAct = X\beta + \beta_{access}*access + u; u = \lambda W u + \epsilon$$

```{r ols_vs_sem}
access_ols <- lm(update(base_formula, physact ~ . + euc_access), data = est_data)
access_sem <- errorsarlm(update(base_formula, physact ~ . + euc_access),
                            data = est_data, listw = Wtr, zero.policy = TRUE)

ols_vs_sem <- list("OLS" = access_ols,  "SEM" = access_sem)
```

@@

sp_hausman(access_ols, access_sem)
```

@@

Okay, given that, the question is whether adding lagged
variables $\gamma W X$ to the model will resolve the discrepancy.

@@

The same basic logic applies to the SLX and SDEM models; if there is no dependence
relationship among the $y$ variables, then the SLX and SDEM will both have unbiased
estimates of $\beta$ and $\gamma$.

```{r access_slx}
access_slx <- lmSLX(update(base_formula, physact ~ . + euc_access), 
                   data = est_data, listw = Wtr, zero.policy = TRUE)
access_sdem <- errorsarlm(update(base_formula, physact ~ . + euc_access),
                          data = est_data, listw = Wtr, zero.policy = TRUE, 
                          Durbin = TRUE)
slx_vs_sdem <- list(
  "SDEM" = access_sdem, 
  "SLX" = access_slx
)
```

As before, we can try a Hausman-style test. In this case the test still suggests
rejecting the null hypothesis, but the parameter esimates are very close to each
other. Additionally, rejecting the null hypothesis would suggest the SDM, which
we rejected on intuitive grounds.

```{r slx_hausman}
slx_hausman <- function(slx, sdem) {
  b1 <- slx$coefficients
  b2 <- sdem$coefficients
  db <- b1 - b2

@@

  tibble(
    Hstat = t(db) %*% MASS::ginv(dV) %*% db, 
    "p-value" = pchisq(Hstat, df = Matrix::rankMatrix(dV), lower.tail = FALSE)
  )
}
slx_hausman(access_slx, access_sdem)
```

@@

```{r lagged_vars}
tr_lagged <- est_data %>%
  st_set_geometry(NULL) %>%
  as_tibble() %>%
  transmute_at(
    vars(density, income, fulltime, college, single, youth, young_adults, seniors, black, asian, hispanic, other, physact),
    ~lag.listw(Wtr, .)
  ) %>%
  mutate(geoid = est_data$geoid) %>%
  dplyr::select(geoid, everything())

names(tr_lagged)[-1] <- str_c("lag.", names(tr_lagged)[-1])
est_data_lagged <- est_data %>% left_join(tr_lagged, by = "geoid")
```

@@

```{r nolag_access}
access_slx <- lm(update(
  base_formula, physact ~ . +  log(lag.density) + log(lag.income) + 
    lag.fulltime + lag.college + lag.single + lag.youth +  lag.young_adults + 
    lag.seniors + lag.black + lag.asian + lag.hispanic + lag.other + euc_access), 
  data = est_data_lagged)
access_sdem <- errorsarlm(update(
  base_formula, physact ~ . +  log(lag.density) + log(lag.income) + 
    lag.fulltime + lag.college + lag.single + lag.youth +  lag.young_adults + 
    lag.seniors + lag.black + lag.asian + lag.hispanic + lag.other + euc_access), 
  data = est_data_lagged, listw = Wtr, zero.policy = TRUE)
```

```{r nolag_access_tests}
slx_hausman(access_slx, access_sdem)
slx1_vs_sdem1 <- list("SLX" = access_slx, "SDEM" = access_sdem)

screenreg(slx1_vs_sdem1, ci.force = TRUE, digits = 4)
```

## Alternative Distances and Aggregations

As discussed on the [Distance](03_distances.html) page, there are four options for
distance in this study:

  - Euclidean distance between tracts and parks
  - Network distance between tracts and parks
  - Euclidean distance between block groups and parks
  - Network distance between block groups and parks

In this section we are going to examine the predictions of our model were we to
change the distance calculation.

### Network vs Euclidean Distance

```{r tr_euc_v_net}
access_sdem_tr_euc <- access_sdem
access_sdem_tr_net <- errorsarlm(update(
  access_sdem, . ~ . - euc_access + net_access), 
  data = est_data_lagged, listw = Wtr, zero.policy = TRUE)

multi_sdem_tr_net <- errorsarlm(update(
  access_sdem, . ~ . - euc_access + net_multi), 
  data = est_data_lagged, listw = Wtr, zero.policy = TRUE)
multi_sdem_tr_euc <- errorsarlm(update(
  access_sdem, . ~ . - euc_access + euc_multi), 
  data = est_data_lagged, listw = Wtr, zero.policy = TRUE)

tr_euc_v_net <- list(
  "Access Euclidean" = access_sdem_tr_euc,
  "Access Network"   = access_sdem_tr_net,
  "Multi Euclidean"  = multi_sdem_tr_euc,
  "Multi Network"    = multi_sdem_tr_net
)
```

The chart below shows a comparison of spatial models of physical activity rates
calculated with Euclidean and network distances. As you can see, the two distance
calculations show virtually the same coefficients and significance across all
borders. The access variable is also virutally identical; the network-based distances
show a slightly smaller mean effect but also a slightly smaller standard 
deviation. The confidence intervals overlap almost entirely.

```{r show_tr_euc_v_net}
texreg::screenreg(tr_euc_v_net, digits = 4)
write_rds(tr_euc_v_net, "data/tr_euc_v_net.rds")
```

Given the challenges in getting network distances for Staten Island --- and the 
immateriality of the effect ---  we would prefer to use a consistent distance
metric across all models.

### Block Group vs Tract

The data available for the study is only available at the tract level. However,
we could down-project the data to the block group to see if that changes any
results.

```{r bg_est_data}
blockgroups <- tigris::block_groups("New York", c("081", "047", "061", "005", "085"), 
                                    class = "sf")

bg_est_data <- blockgroups %>%
  transmute(geoid = GEOID, tract = substr(geoid, 1, 11)) %>%
  left_join(read_rds("data/tracts.rds") %>% st_set_geometry(NULL), by = c("tract" = "geoid") ) %>%
  filter(!is.na(borough)) %>%
  left_join(read_rds("data/bg_logsums.rds"))

Wbg <- nb2listw(poly2nb(bg_est_data), style = "W", zero.policy = TRUE)

bg_lagged <- bg_est_data %>%
  st_set_geometry(NULL) %>%
  transmute_at(
    vars(density, income, fulltime, college, single, youth, young_adults, seniors, black, asian, hispanic, other, physact),
    ~lag.listw(Wbg, ., zero.policy = TRUE)
  ) %>%
  mutate(geoid = bg_est_data$geoid) %>%
  dplyr::select(geoid, everything())

names(bg_lagged)[-1] <- str_c("lag.", names(bg_lagged)[-1])
bg_est_data_lagged <- bg_est_data %>% left_join(bg_lagged, by = "geoid")
```

We are unable to estimate some of the models because of collinearity issues;
this is a problem in the spatially lagged predictor variables especially.
But we can look at what happens when we make block groups the unit of analysis
in the SEM.

```{r bg_models}
access_sem_bg <- errorsarlm(update(base_formula, physact ~ . + euc_access), 
  data = bg_est_data_lagged, listw = Wbg, zero.policy = TRUE)
bg_models <- list("Tracts" = access_sem, "BG" = access_sem_bg) %>% write_rds("data/bg_models.rds")
texreg::screenreg(bg_models, digits = 4)
```

@@

```{r estimate_sdems}
base_sdem <- update(access_sdem, formula = . ~ . - euc_access,
                    data = est_data_lagged, listw = Wtr, zero.policy = TRUE)

# estimate models
pa_access_sdem <- access_sdem
pa_multi_sdem  <- update(base_sdem, .~ . + euc_multi)
pa_walk10_sdem <- update(base_sdem, .~ . + walk10)
```

@@

```{r access_sdm}
pa_models <- list("Base"        = base_sdem,
                  "Logsum"      = pa_access_sdem,
                  "Attributes"  = pa_multi_sdem,
                  "10-min walk" = pa_walk10_sdem)
```
